\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-05-01 archive.tex
better dreams for towers

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit, procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can quickly categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of problems can often be described as
some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model approaches a specific problem by searching for a program solving
it. It gradually grows its explicit domain-specific knowledge for a
class of problems by assembling a library of code containing concepts
useful for the domain, and by training a neural network to quickly
infer, for a specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous to the
explicit concepts and implicit procedural skill that human experts
develop.  The model is an instance of what in the machine learning
community is called a wake-sleep algorithm, where for us, `waking'
corresponds to solving problems, while a pair of `sleep' cycles
correspond to improving the library of explicit concepts and training
the neural net to improve implicit procedural knowledge of how to
write code.

@@

Our model iteratively creates new library routines that build on
concepts acquired earlier in its learning trajectory, growing a library with nested
hierarchies of code.  We think of this cumulative nesting of
abstractions as a variety of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the library). For example, the model
learns to sort sequences of numbers by invoking a library component 4
layers deep, or draws the leftmost pairs of images in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to our approach as an instance of `deep program learning'.

\begin{figure}
  \includegraphics[width = \textwidth]{figures/DeepLogo_4_15}
  \includegraphics[width = \textwidth]{figures/DeepList_4_17}
  \caption{Model solves tasks (left; top, task is to draw an image. bottom, task is input/output mapping) by writing programs (right column) where these programs invoke explicit declarative knowledge in the form of library routines (middle) jointly learned w/ the solutions. Library code can call each other (arrows), forming a layered network of domain-specific concepts. Fantasies (below library) are tasks sampled by randomly combining code from library. Fantasies are used to train a neural net to help search for programs.}\label{exampleDSL}
\end{figure}

@@

Mathematically this compression takes the form of
finding the library maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We replace this marginal with an AIC approximation~\cite{akaike1998information}
and marginalize over refactorings of programs found during waking, minimizing the following expression,
which can be interpreted as a kind of compression:
\begin{equation}
\underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
  X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
\end{equation}
But a program has infinitely many possible refactorings, rendering
Eq.~\ref{AIC} intractable.  Rather than consider every refactoring we
bound the number of $\lambda$-calculus evaluation steps separating a
program from its refactoring.  Now the number of refactorings is
finite but astronomically large: Figure~\ref{mapFactor}A diagrams a
problem where the agent rediscovers the higher-order function
\code{map} starting from the basics of Lisp and the Y-combinator, but
where there are approximately $10^{14}$ possible refactorings -- a quantity
that grows exponentially both as a function of program size and a
function of the bound on evaluation steps. How can we tame this
combinatorial explosion?

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system typically composes around 20 new library routines, and discovered multiple
higher-order functions. Each round of abstraction built on
concepts discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{exampleDSL}).
This incremental, modular learning of deep hierarchies of library
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the abstraction phase
of sleep).

\subsubsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that synthesize text editing programs ship in Microsoft
Excel~\cite{gulwani2011automatin}.  This prior work uses
hand-engineered libraries of primitives and hand-engineered search strategies.  Here, we
will show that we can jointly learn both these ingredients and surpass
the state-of-the-art domain-general program synthesizers on a standard
text editing benchmark.

%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

We trained our system on 128 automatically-generated text editing tasks, with 4 input/output examples each.
We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems within 10 minutes with an average search time of 235 seconds.
After learning,
it solves 79.6\%, and does so much faster,
solving them in an average of 40 seconds.
As of the 2017 SyGuS competition,
the best-performing synthesizer (CVC4) solves 82.4\% of the problems ---
but here, the competition conditions are 1 hour \& 8 CPUs per problem,
and with this more generous compute budget we
surpass
this previous
result and solve
84.3\% of the problems.
SyGuS additionally comes with a
different hand-engineered libraries of primitives \emph{for each text editing problem}. %\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task. For these experiments, our system did not use this assistance.}
Here  we learned a single library of text-editing concepts
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

    \\\bottomrule \\\\
    \multicolumn{2}{c}{
      \begin{tabular}{ccc}
        Dreams before learning&Dreams after learning&\\
        \begin{tabular}{c}
          \includegraphics[width = 0.4\textwidth]{figures/initialDreams/montage.png}
        \end{tabular}&
        \begin{tabular}{c}
          \includegraphics[width = 0.4\textwidth]{figures/finalDreams/cherry_random_montage.png}
        \end{tabular}&
        \rotatebox[origin=c]{90}{Cherry picked\hspace{2.5cm}Random}
      \end{tabular}
    }
  \end{tabular}
  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: Example learned library routines. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input. \textbf{Bottom}: Sixteen dreams, or sampled programs, from library before (left) and after (right) learning. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoMaster}
  \end{figure}

@@

Inspired by the classic AI `copy demo' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{tower}, top).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.
The system starts with the same control flow primitives as with
LOGO graphics,
and inside its learned library we find parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
including latent concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \centering  \begin{tabular}{cc}
    \toprule
    \multicolumn{2}{c}{Subset of 112 Tasks}\\\midrule
    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower_montage_10.png}}\\\midrule \\
    \toprule Example learned Library components&Samples written using final library\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 6cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Top}: Ten (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Bottom left}: Three learned library routines. These components act like parametric options~\cite{stolle2002learning},
    giving higher-level building blocks that the agent can use to plan. \textbf{Bottom Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
\end{figure}

@@

In solving these tasks, \system first discovers the building blocks of vector algebra,
like inner products, vector sums, and norms.
It then uses this vocabulary to infer library routines for common
physical relationships like the vector form inverse square laws
and integrating second derivatives over time,
effectively undergoing a `change of basis' from the initial Lisp-style primitives into a
physics-style basis. Figure~\ref{librariesToLanguages}~(left) illustrates a subset
of the provided tasks and learned library for this domain.

@@

        %% Center-of-mass:\\
        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
        \hspace{0.2cm}\blueCode{($f_{10}$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        %% Integrating angle over time:\\
        %% $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
        %% \code{1, 3, 0, 2$\to$7}\\
        %% \code{-5, 0, 10, 1$\to$0}\\
        %% \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
        Newton's Second Law:\\
        $m$, $\{\vec{F}_i \}\mapsto \frac{1}{m}\sum_i\vec{F}_i = a$\\
        \code{2, [[2 4 1]]$\to$[1 2 0.5]}\\
        \code{3, [[0 0 2] [3 0 1]]$\to$[1 0 0.333]}\\
        \blueCode{$f(m$, forces$) = $($f_4$ ($f_7$ m) ($f_3$ forces))}
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

@@

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/regex_marginal_test_unigram_gen.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/regex_task_train_unigram_gen.png}
      \end{tabular}
  \end{tabular}
  \caption{Test set performance across wake/sleep iterations. Each curve is a run with a different random seed. Teal: Full model. Orange: Dreaming only (no library learning). Purple: Consolidation only (no recognition model). Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks). Generative text modeling plots show (top) posterior predictive likelihood of held-out strings on held out tasks and (bottom) lower bound on marginal likelihood of held out tasks, both of which are averaged per character.}\label{learningCurves}      
\end{figure}

@@

Our goal was to build a model capturing
humanlike aspects of learning to be an expert in a new domain of
problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
representation of a domain's conceptual structure,
while learning to quickly combine these concepts into programs
solving specific problems within the domain.
Two main findings of our work are that, first, a single
program-induction system can learn to solve large sets of problems
from many qualitatively different domains; and second, that fully
acquiring expertise in these domains hinges both upon explicit
declarative knowledge, and implicit procedural skill.  We have
deliberately placed the interplay of declarative and procedural
knowledge at the center of our model, and believe that both these
kinds of learning are crucial for building agents that, like humans,
autonomously learn to navigate a new domain of problems.

\system builds on several generations of research in AI, program synthesis,
and cognitive science, with program induction being one of the oldest
theoretical frameworks for concept learning within artificial
intelligence~\cite{solomonoff1964formal}, and the conceptually allied
`Language of Thought Hypothesis' being almost as
old~\cite{fodor1975language}. Recent neural program synthesis systems
pair a fixed domain-specific programming language (a `DSL') to a
learned neural network that guides program
search~\cite{DBLP:journals/corr/abs-1804-01118,balog2016deepcoder,devlin2017neural}, while recent
symbolic AI research has developed frameworks for learning the DSL by
inferring reusable pieces of
code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
A main goal of this work is to combine and refine these techniques
with the intention of building agents that, like humans, develop
domain expertise for new classes of problems.  Work in cognitive
modeling has often exploited program-like representations for both
concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
learning~\cite{logical,Ullman2012},
and we view these models of concept learning as
a kind of program induction,
and see theory learning as similar to inducing a DSL or library.

@@

Our model's domain knowledge grows over a series of wake/sleep cycles,
with the solutions to easier problems bootstrapping
solving of harder tasks.
This dynamic is related to, but distinct from, curriculum learning. In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
But humans select their tasks in much richer ways,
and can even generate their own tasks to solve,
either as stepping stones toward hard, unsolved problems or motivated by considerations like curiosity and aesthetics.
Building agents that generate their own problems in these humanlike ways is a necessary next step
if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
both of which go far beyond the relatively small-scale demonstrations in this work.

>>>>

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit, procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can, at a glance, categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of problems can often be described as
some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model approaches a specific problem by searching for a program solving
it. It gradually grows its domain-specific knowledge for a
class of problems by assembling a library of code containing concepts
useful for the domain, while jointly training a neural network to quickly
infer, for a specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous to the
explicit concepts and implicit procedural skill that human experts
develop.  The model is an instance of what in the machine learning
community is called a wake-sleep algorithm, where for us, `waking'
corresponds to solving problems, while a pair of `sleep' cycles
correspond to improving the library of explicit concepts and training
the neural net to improve implicit procedural knowledge of how to
write code.

@@

Our model iteratively creates new library routines that build on
concepts acquired earlier in its learning trajectory, growing a library with nested
hierarchies of code.  We think of this cumulative nesting of
abstractions as a variety of deep representation
learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
subset of these learned networks (the library). For example, the model
learns to sort sequences of numbers by invoking a library component 4
layers deep, or draws the topmost image in
Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
refer to our approach as an instance of `deep program learning'.

\begin{figure}\centering
  \includegraphics[width = \textwidth]{figures/deepLogo_may1}
  \includegraphics[width = \textwidth]{figures/deepList_may1}
  \caption{Model solves tasks (left; top, task is to draw an image. bottom, task is input/output mapping) by writing programs (right column) where these programs invoke explicit declarative knowledge in the form of library routines (middle) jointly learned w/ the solutions. Library code can call each other (arrows), forming a layered network of domain-specific concepts. %% Fantasies (below library) are tasks sampled by randomly combining code from library. Fantasies are used to train a neural net to help search for programs.
  }\label{exampleDSL}
\end{figure}

@@

Mathematically this compression takes the form of
finding the library maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
We %% replace this marginal with an AIC approximation~\cite{akaike1998information}
%% and
marginalize over refactorings of programs found during waking, seeking to minimize the following expression,
which can be interpreted as a kind of compression:
%% \begin{equation}
%% \underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
%% \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
%%   X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
%% \label{AIC}
%% \end{equation}
\begin{equation}
\underbrace{  -\log \probability[\mathcal{D}]_{\substack{\phantom{p}\\\\\\\\\\}}}_{\text{Description length of library}}\underbrace{-\log \int \probability[\theta|\mathcal{D}]\prod_{x\in X}\sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]\;\mathrm{d}\theta}_{\text{Description length of tasks given library}}
%% \underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
%% \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
%%   X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
\label{AIC}
\end{equation}
But a program has infinitely many possible refactorings, rendering
Eq.~\ref{AIC} intractable.  Rather than consider every refactoring we
bound the number of $\lambda$-calculus evaluation steps separating a
program from its refactoring.  Now the number of refactorings is
finite but astronomically large: Figure~\ref{mapFactor}A diagrams a
problem where the agent rediscovers the higher-order function
\code{map} starting from the basics of Lisp and the Y-combinator, but
where there are approximately $10^{14}$ possible refactorings -- a quantity
that grows exponentially both as a function of program size and a
function of the bound on evaluation steps. How can we tame this
combinatorial explosion?

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system typically composes around 20 new library routines, and rediscovers
higher-order functions such as \code{filter}. Each round of abstraction built on
concepts discovered in earlier sleep cycles --- for example the
agent first learns \code{filter}, then uses
it to learn to take the maximum element of a list, then
uses that routine to learn a new library routine for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{exampleDSL}).
This incremental, modular learning of deep hierarchies of library
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the abstraction phase
of sleep).

\subsubsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that synthesize text editing programs ship in Microsoft
Excel~\cite{gulwani2011automating}.  This prior work uses
hand-engineered libraries of primitives and hand-engineered search strategies.  Here, we
will show that we can jointly learn both these ingredients and perform comparably to
the state-of-the-art domain-general program synthesizers on a standard
text editing benchmark.

We trained our system on 128 automatically-generated text editing tasks, with 4 input/output examples each.
We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Figure~\ref{textEditExamples} shows representative text-editing problems. Before any learning,
\system solves 3.7\% of the problems within 10 minutes with an average search time of 235 seconds.
After learning,
it solves 79.6\%, and does so much faster,
solving them in an average of 40 seconds.
As of the 2017 SyGuS competition,
the best-performing synthesizer (CVC4) solves 82.4\% of the problems ---
but here, the competition conditions are 1 hour \& 8 CPUs per problem,
and with this more generous compute budget we
surpass
this previous
result and solve
84.3\% of the problems.
SyGuS additionally comes with a
different hand-engineered libraries of primitives \emph{for each text editing problem}. %\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task. For these experiments, our system did not use this assistance.}
Here  we learned a single library of text-editing concepts
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.
\begin{figure}[h]\centering
  \begin{tabular}{ccc}
    \fbox{\begin{tabular}{c}
        \;\,Dream Coder$\to$(Coder)\\
        Domain Expertise$\to$(Expertise)
    \end{tabular}}&
    \fbox{\begin{tabular}{c}
        Temple Anna H$\to $Dr. TAH\\
        Lara Gregori$\to $Dr. LG
    \end{tabular}}&
    \fbox{\begin{tabular}{c}
      this test-split here.$\to$split here\\
      another-text.test for you$\to$text
    \end{tabular}}
  \end{tabular}
  \caption{Three example text editing tasks. Each task is specified by input/output pairs (notated above as input$\to$output), and the goal is to infer a program producing the input/output mapping.}\label{texEditProblems}
\end{figure}

@@

    \\\bottomrule \\\\
    \multicolumn{2}{c}{
      \begin{tabular}{ccc}
        Dreams before learning&Dreams after learning&Dreams after learning (cherry picked)\\
          \includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}&
          \includegraphics[width = 0.3\textwidth]{figures/dreams/final25}&
          \includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}
      \end{tabular}
    }
  \end{tabular}
  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: Example learned library routines. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input. \textbf{Bottom}: 25 dreams, or sampled programs, from library before (left column) and after (middle \& right) learning. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoMaster}
  \end{figure}

@@

Inspired by the classic AI `copy demo' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{tower}, top).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.
The system starts with the same control flow primitives as with
LOGO graphics,
and inside its learned library we find parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
including latent concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \centering  \begin{tabular}{cc}
    \toprule
    \multicolumn{2}{c}{Subset of 112 Tasks}\\\midrule
    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower_montage_10.png}}\\\midrule \\
    \toprule Example learned library components&Dreams from learned library\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 8cm]{figures/tower/dreams/cherry_montage.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Top}: Ten (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Bottom left}: Three learned library routines. These components act like parametric options~\cite{stolle2002learning},
    giving higher-level building blocks that the agent can use to plan. \textbf{Bottom Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
\end{figure}

@@

In solving these tasks, \system first discovers the building blocks of vector algebra,
like inner products, vector sums, and norms.
It then uses this vocabulary to infer library routines for common
physical relationships like the vector form of inverse square laws
and integrating second derivatives over time,
effectively undergoing a `change of basis' from the initial Lisp-style primitives into a
physics-style basis. Figure~\ref{librariesToLanguages}~(left) illustrates a subset
of the provided tasks and learned library for this domain.

@@

        %% Center-of-mass:\\
        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
        \hspace{0.2cm}\blueCode{($f_{10}$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        %% Integrating angle over time:\\
        %% $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
        %% \code{1, 3, 0, 2$\to$7}\\
        %% \code{-5, 0, 10, 1$\to$0}\\
        %% \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
        Newton's Second Law:\\
        $m$, $\{\vec{F}_i \}\mapsto \frac{1}{m}\sum_i\vec{F}_i = a$\\
        \code{2, [[2 4 1]]$\to$[1 2 0.5]}\\
        \code{3, [[0 0 2] [3 0 1]]$\to$[1 0 0.333]}\\
        \blueCode{$f(m$, forces$) = $($f_4$ ($f_7$ m) ($f_3$ forces))}
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

@@

    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/regex_marginal_test_unigram_gen.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/regex_task_train_unigram_gen.png}
      \end{tabular}
  \end{tabular}
  \caption{Test set performance across wake/sleep iterations. Each curve is a run with a different random seed. Teal: Full model. Orange: Dreaming only (no library learning). Purple: Abstraction only (no recognition model). Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks). Generative text modeling plots show (top) posterior predictive likelihood of held-out strings on held out tasks and (bottom) lower bound on marginal likelihood of held out tasks, both of which are averaged per character.}\label{learningCurves}      
\end{figure}

@@

Our goal was to build a computational model capturing
aspects of learning to be an expert in a new domain of
problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
representation of a domain's conceptual structure,
while learning to quickly combine these concepts into programs
solving specific problems within the domain.
Two main findings of our work are that, first, a single
program-induction system can learn to solve large sets of problems
from many qualitatively different domains; and second, that fully
acquiring expertise in these domains hinges both upon explicit
declarative knowledge, and implicit procedural skill.  We have
deliberately placed the interplay of declarative and procedural
knowledge at the center of our model, and believe that both these
kinds of learning are crucial for building agents that, like humans,
autonomously learn to solve a new domain of problems.

\system builds on several generations of research in AI, program synthesis,
and cognitive science, with program induction being one of the oldest
theoretical frameworks for concept learning within artificial
intelligence~\cite{solomonoff1964formal}, and the conceptually allied
`Language of Thought Hypothesis' being almost as
old~\cite{fodor1975language}. Recent neural program synthesis systems
pair a fixed domain-specific programming language (a `DSL') to a
learned neural network that guides program
search~\cite{DBLP:journals/corr/abs-1804-01118,balog2016deepcoder,devlin2017neural}, while recent
symbolic AI research has developed frameworks for learning the DSL by
inferring reusable pieces of
code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
A main goal of this work is to combine and refine these techniques
with the intention of building agents that develop
domain expertise for new classes of problems.  Work in cognitive
modeling has often exploited program-like representations for both
concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
learning~\cite{logical,Ullman2012},
and we view these models of concept learning as
a kind of program induction,
and see theory learning as similar to inducing a DSL or library.

@@

Our model's domain knowledge grows over a series of wake/sleep cycles,
with the solutions to easier problems bootstrapping
solving of harder tasks.
This dynamic is related to, but distinct from, curriculum learning. In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks, but here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
But humans select their tasks in much richer ways,
and can even generate their own tasks to solve,
either as stepping stones toward hard, unsolved problems or motivated by considerations like curiosity and aesthetics.
Building agents that generate their own problems in these humanlike ways is a necessary next step
if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
both of which go far beyond the relatively small-scale demonstrations in this work.
